\chapter*{Acknowledgments}
\addcontentsline{toc}{chapter}{Acknowledgments}
\setheader{Acknowledgments}

Whilst this thesis bears my name as the only author, all of this work and my whole personal and
professional development would not have been remotely possible without the help of a number of
people. There is no way a few paragraphs can express my gratitude satisfactorily, but I want you to
know that your support has been invaluable and essential.

My two promotors Stephanie and Przemek have been my guiding light throughout this four-year
scientific journey. Ronald has provided a cutting-edge testbed to enable groundbreaking experiments.
The committee members Arie, Carmina, Lydia and Rodney have provided solid feedback to help me refine
this dissertation and have accepted to wear Medieval heat-trapping robes at the end of June to
attend my defence. My fellow group mates Álvaro, Bart, Bethany, Francisco, Guus, Hana, Janice, Ravi,
Scarlett, Soubhadra and Thomas, and other present and past TU Delfters Alejandro, Aram, Axel, David,
David, Eric, Gayane, Guilherme, Helena, Ingmar, Jasper, Josh, Joël, Kanchan, Kenneth, Luise,
Mariagrazia, Matt, Matt, Matteo, Nico, Nico, Önder, Paul, Ravi, Remon, Siddhant, Sébastian, Thom,
Thomas, Tim, Vicky, Vito, Wojtek and Yves have been the pillar of my work and life at TU Delft. My
friends from Italy Alessio, Davide, Davide, Edoardo, Eleonora, Filippo, Francesco, Giulia, Manuela,
Pierluigi, Rosita and Simone, and those found in the Netherlands Costas, Dimitris and Patrick have
never failed to make me feel loved and never cease to being a massive source of joy, energy and
inspiration. My whole family, and particularly Mamma, Papà, Luigi and my companion Giulia are my
shell, my den, my all-inclusive five-star suite, they are me, and I am them.

\begin{flushright}
{\itshape
Carlo \\
Delft, June 2023
}
\end{flushright}
