\begin{center}

{\Large\titlefont\bfseries Propositions}

\medskip

accompanying the dissertation

\medskip

% Print the title.
{\makeatletter
\titlestyle\bfseries\large\@title
\makeatother}

% Print the optional subtitle.
{\makeatletter
\ifx\@subtitle\undefined\else
\titlefont\titleshape\@subtitle
\fi
\makeatother}

\medskip

by

\medskip

% Print the full name of the author.
\makeatletter
{\large\titlefont\bfseries\@firstname\ {\titleshape\@lastname}}
\makeatother

\end{center}

\bigskip

\begin{enumerate}
    \item Quantum networks can only become a commercially viable product if the nodes thereof can be
          programmed with platform-independent software.
    \item The overhead incurred by software abstractions for quantum network nodes does not affect
          the ability of said nodes to establish entanglement with one another.
    \item Multi-processing operating systems are not only beneficial, but also essential, for the
          operability of multi-user quantum networks.
    \item Packet switching routing schemes will eventually outclass path reservation-based resource
          allocation methods in quantum networks.
    \item The scientific integrity of a doctoral candidate must never be assumed, and should
          instead be thoroughly trained and assessed by an independent committee.
    % \item An academic community that values the volume of publications more than their quality
    %       undermines the integrity of its members.
    \item Magritte's \emph{pipe} is, in fact, a pipe.
    \item Progress in various areas of technology, including telecommunications, is to a large
          extent to be attributed to many people's addiction to porn.
    \item The \emph{National Meatless Week} is an unintentionally fallacious campaign.
    \item Any definition of what constitutes a work of art on the basis of (subjective) beauty,
          market value, technique, medium, the artist's status, or even the artist's sentience, is
          bound to become obsolete some time later in history.
    \item At the time of writing, one can either be a feminist or a sexist.
\end{enumerate}

\bigskip
\noindent
Propositions 1 to 3 pertain to this thesis.

\bigskip
\bigskip

% The following text is dictated by the promotieregelement.
\begin{center}
These propositions are regarded as opposable and defendable, and have been approved as such by the
promotors prof.\ dr.\ S.\ D.\ C.\ Wehner and dr.\ P.\ Pawełczak.
\end{center}
